---
title: "Creating and Analyzing a Simple Map with dodgr"
output: html_document
---

# Vignette: Creating and Analyzing a Simple Map with `dodgr`

This vignette demonstrates how to create a simple map using the `dodgr` package, add points to it, and analyze the resulting network. We will explore how to create a line, add points, and calculate paths using the `dodgr` package.

## 1. Setup

First, we load the necessary libraries and the `dodgr` package:

```{r}
library(sf)
devtools::load_all(here::here("../dodgr/"))
devtools::load_all(here::here("."))
library(ggplot2)
library(ggrepel)
library(dplyr)
dodgr_cache_off()
```

## 2. Creating a Simple Road

We begin by creating a simple straight road represented as a line string:

```{r}
road <- st_linestring(matrix(c(0,0, 10,0), ncol=2, byrow=TRUE))
roads <- st_sf(
  geometry = st_sfc(road, crs = 4326),
  highway = "residential",
  id = 1,
  artificial = FALSE
)

```

### Comment:
This creates a single road segment with a specified coordinate reference system (CRS).

## 3. Adding Points

Next, we add two points near the road but not directly on it:

```{r}
point1 <- st_point(c(2, 1))
point2 <- st_point(c(8, -1))
points <- tibble(x=c(2,8), y=c(1,-1), id=c(2,3)) %>%
  sf::st_as_sf(coords=c("x", "y"), remove=FALSE)
```

### Comment:
The points are converted to an `sf` object, allowing spatial operations.

## 4. Creating a Network

We create an initial network from the road:

```{r}
net_initial <- weight_streetnet(roads, id_col = "id")
v_initial <- dodgr_vertices(net_initial)
```

### Comment:
This function weights the street network, which is essential for routing and pathfinding.


## 6. Connect points to network

```{r}
test_comp <- function(x) dodgr::dodgr_components(x%>%rename(cmp_old=component))%>%count(cmp_old, component)

net_connected <- add_verts_to_graph(net_initial, st_coordinates(points))

ggplot(net_connected)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat))+
  geom_point(aes(x = from_lon, y=from_lat), color="blue")+
  geom_point(aes(x = x, y=y), color="red", data=points)
```



```{r}
paths_e <- dodgr_paths(net_connected, from = st_coordinates(points)[1,], to = st_coordinates(points)[2,], vertices = FALSE)
message("Paths results:")
print(paths_e)
ggplot(net_connected)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat))+
  geom_point(aes(x = from_lon, y=from_lat), color="blue")+
  geom_point(aes(x = x, y=y), color="red", data=points)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat), data=net_connected[unlist(paths_e),], color="pink")

```



```{r}
curved_road <- st_linestring(matrix(c(
  0, 0,
  2, 1,
  4, 1,
  6, 0,
  8, 0
), ncol=2, byrow=TRUE))

crossing_road <- st_linestring(matrix(c(
  4, -1,
  4, 2
), ncol=2, byrow=TRUE))

disconnected_road <- st_linestring(matrix(c(0,-2, 8,-2), ncol=2, byrow=TRUE))

roads_sf <- st_sf(
  geometry = st_sfc(curved_road,
                    crossing_road,
                    disconnected_road, crs=4326),
  highway = c("residential", "residential", "residential"),
  id = c(1,
         2,
         3)
)
roads_sf <- roads_sf[c(1,3),]

ggplot(data=roads_sf)+geom_sf()

```


# Create points that should connect to different parts

```{r}

points_sf <- st_sf(
  geometry = st_sfc(
    st_point(c(0, -3)),    # Should connect to disconnected road
    st_point(c(2, 0)),    # Should connect to curved road
    st_point(c(4, 0.5)),  # Should connect to both roads
    st_point(c(6, 1)),    # Should connect to curved road
    crs = 4326
  ),
  id = c(3, 4, 5,6)
)
points_sf$x <- st_coordinates(points_sf)[,1]
points_sf$y <- st_coordinates(points_sf)[,2]
ggplot(data=roads_sf)+geom_sf() + geom_sf(data=points_sf)
```


# Create initial dodgr network
```{r}

net_initial <- weight_streetnet(roads_sf, id_col = "id")
v_initial <- dodgr_vertices(net_initial)
message("\nInitial network structure:")
message("- Number of vertices: ", nrow(v_initial))
message("- Number of edges: ", nrow(net_initial))

```


```{r}
v2 <- v_initial%>%filter(component==2)%>%
  st_as_sf(coords=c("x", "y"), remove=FALSE)%>%
  slice(1)
net_b_1 <- net_initial%>%filter(component==1)
net_b_2 <- net_initial%>%filter(component==2)
net_b_12 <- add_verts_to_graph(net_b_1, v2)%>%
  add_edges_to_graph(v2, wt_profile="motorcar", highway = "residential")%>%
  bind_rows(net_b_2%>%mutate(edge_id=as.character(edge_id)))%>%
  std_graph()

test_comp(net_b_12)

```


```{r}
ggplot(net_initial) +
  geom_sf(data=roads_sf, color="gray")+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat), color="pink", alpha=1/2)+
  geom_label_repel(aes(x = from_lon, y=from_lat, label=from_id),data=net_b_12%>%distinct(from_lon, from_lat, from_id))+
  geom_point(aes(x = x, y=y), color="red", data=points_sf)

```


```{r}
net_b <- net_b_12#%>%break_long_edges(max_d = 400000)
```


```{r}
initial_paths <- try(dodgr_paths(net_b, from = points_sf%>%st_coordinates(), vertices = FALSE))
message("\nInitial path attempt result:")
ggplot(net_b)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat))+
  geom_point(aes(x = from_lon, y=from_lat), color="blue")+
  geom_point(aes(x = x, y=y), color="red", data=points_sf)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat), data=net_b[unlist(initial_paths),], color="pink")

```




```{r}
p2 <- st_coordinates(points_sf)
net_connected2 <- add_verts_to_graph(net_b,p2, bidirectional = FALSE)%>%
  add_edges_to_graph(p2, wt_profile="motorcar", highway = "residential")%>%
  std_graph()

dodgr::dodgr_components(net_connected2%>%select(-component))%>%count(component)
```


# 6. Calculate path
```{r}
paths_e <- dodgr_paths(net_connected2, from = st_coordinates(points_sf%>%filter(id==6)), to = st_coordinates(points_sf%>%filter(id==3)), vertices = FALSE)
message("Paths results:")
print(paths_e)
v_connected2 <- dodgr_vertices(net_connected2)
ggplot(net_connected2)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat, color=factor(component)))+
  geom_point(aes(x = x, y=y), data=v_connected2)+
  geom_label(aes(x = x, y=y, label=id), color="black", data=points_sf)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat), data=net_connected2[unlist(paths_e),], color="orange", linewidth=3, alpha=1/3) +
  geom_sf(data=roads_sf, color="purple")
```

## create bidirectional edge splits


```{r}
net_connected2b <- add_verts_to_graph(net_b,p2, bidirectional = TRUE)%>%
  add_edges_to_graph(p2, wt_profile="motorcar", highway = "residential")%>%
  std_graph()

dodgr::dodgr_components(net_connected2b%>%select(-component))%>%count(component)
```


# 6. Calculate path
```{r}
paths_eb <- dodgr_paths(net_connected2b, from = st_coordinates(points_sf%>%filter(id==6)), to = st_coordinates(points_sf%>%filter(id==3)), vertices = FALSE)
message("Paths results:")
print(paths_eb)
v_connected2b <- dodgr_vertices(net_connected2b)
ggplot(net_connected2b)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat, color=factor(component)))+
  geom_point(aes(x = x, y=y), data=v_connected2b)+
  geom_label(aes(x = x, y=y, label=id), color="black", data=points_sf)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat), data=net_connected2b[unlist(paths_eb),], color="orange", linewidth=3, alpha=1/3) +
  geom_sf(data=roads_sf, color="purple")
```


## Use hampi data

```{r}
graph <- weight_streetnet (hampi, wt_profile = "foot")
table(graph$component)
dim (graph)
verts <- dodgr_vertices (graph)
```

## connect components 1 and 2 

```{r}
v2 <- verts%>%filter(component==2)%>%
  st_as_sf(coords=c("x", "y"), remove=FALSE)%>%
  slice(1)
net_b_1 <- graph%>%filter(component==1)
net_b_2 <- graph%>%filter(component==2)
net_b_o <- graph%>%filter(!component%in%c(1:2))
net_b_12 <- add_verts_to_graph(net_b_1, v2)%>%
  add_edges_to_graph(v2, wt_profile="motorcar", highway = "residential")%>%
  bind_rows(net_b_2%>%mutate(edge_id=as.character(edge_id)))%>%
  bind_rows(net_b_o%>%mutate(edge_id=as.character(edge_id)))%>%
  std_graph()
table(net_b_12$component)
```


```{r}
set.seed (2)
npts <- 100
xy <- data.frame (
    x = min (verts$x) + runif (npts) * diff (range (verts$x)),
    y = min (verts$y) + runif (npts) * diff (range (verts$y))
)
```


## using dodgr
```{r}
graph_dodgr <- add_nodes_to_graph (graph, xy)
dim (graph_dodgr) # more edges than original
```




## find xy points connecting to the same edge


```{r}
pt_matches <- match_pts_to_graph(graph, xy, distances = TRUE)
pt_max_edge <- pt_matches$index%>%table%>%sort%>%tail(1)
pt_1 <- pt_matches%>%filter(index==names(pt_max_edge))%>%
  mutate(component=graph$component[index])%>%
  filter(component==1)%>%
  arrange(x,y)
stopifnot(nrow(pt_1)>1)


```

## using dodgr
```{r}
pt_path <- dodgr_paths(graph = graph, 
            from = head(pt_1[,c("x", "y")],1),
            to = tail(pt_1[,c("x", "y")],1), vertices = FALSE)
print(pt_path)



```


## using dodgrconnect

```{r}
graph_connect <- add_verts_to_graph(graph, xy)%>%
  add_edges_to_graph(xy, wt_profile="foot", highway = "residential")%>%
  std_graph()
pt_path <- dodgr_paths(graph = graph_connect, 
            from = head(pt_1[,c("x", "y")],1),
            to = tail(pt_1[,c("x", "y")],1), vertices = FALSE)
print(pt_path)

```


```{r}
ggplot(graph_connect)+
  #geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat, color=factor(component)))+
  #geom_point(aes(x=x, y=y), data=pt_1[c(1,nrow(pt_1)),], color="blue", size=4)+
  #geom_point(aes(x=x, y=y), data=xy)+
  geom_segment(aes(x = from_lon, y=from_lat, xend = to_lon, yend = to_lat),
data=graph_connect[unlist(pt_path),], color="orange", linewidth=3, alpha=1/3)#+geom_sf(data=hampi, color="purple")

```

---




This vignette provides a step-by-step guide to creating and analyzing a simple map using `dodgr`. If you have any specific issues or need further clarification, feel free to ask!
